MSF_Groupwork-NBA_shotlog

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
final changes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```
#####Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
#####Date submitted: 16/10/2016

***

##Introduction

@@

** TO GET UPDATED AS A FINAL STEP**

**1. Team related**  
  a. Home Advantage analysis  
  b. Effect of rest on teams' performance

**2. Player related**  
  a. Effects on Accuracy  
  b. Shot Clock Pressure  
  c. Fatigue Effect (Quarter Accuracy Rate)    
  d. Game result impact by field goal attempts  
  e. Defender Proximity Analysis  
  f. Shooting Distance vs Shot Accuracy   
  g. Hot Hand Theory

###Dataset Description:

@@

###Dataset Strengths

@@

###Dataset Limitations  
  There are several limitations of the dataset:  
    1. The dataset consists only 1 year NBA season stats (2014-2015)  
    2. The dataset comprise only the data for regular season matches  
    3. The playoff and other matches during the year are not included
    4. The dataset does not include information on free throw shots
    5. The dataset does not include final scores of the games

@@

## Home Advantage analysis
###Introduction

@@

```{r, echo=FALSE}
#plot showing home and away win ratio per team
ggplot(advdf, aes(x = advdf$WINNER)) + geom_bar(aes()) + ylab("game count") + xlab("winner")
```

@@

```{r, echo=FALSE}
#plot showing home and away win ratio per team
ggplot(advdf3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "wins HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "wins  AWAY")) + ylab("Win ratio") + xlab("Team Names")
```

@@

***

##Effect of rest on teams' performance

@@

###Findings

@@

***

##Effects on accuracy

@@

graph2 <- ggplot(NULL) + geom_smooth(data = df.acc.secondsplayers3, aes(x = seconds.left, y = acc.top, color = "Top Players")) + geom_smooth(data = acc, aes(x = seconds.left, y = accuracy, color = "Average"))
graph2 + labs(title = "Comparison - Accuracy per second on shot clock left for Top Players vs Overall Average") + xlab("Seconds left on shot clock") + ylab("Accuracy")

@@

***

## Game result impact by field goal attempt

@@

### Impact of Top Scorers

@@

```
**Table 9**: Result for the logistic regression analysis for FGA and FG% of 15 players

@@

**Table 10**: Result for the anova for FGA and FG% for 15 players

@@

**Table 11**: Result for the logistic regression analysis for FGA and FG% of the 3 special players

@@

**Table 12**: Result for the anova for FGA and FG% for the 3 special players

@@

##Defender Proximity Analysis  
###Closest Defender Distance vs Shot Accuracy
####Introduction  
This analysis focuses on whether there is a significant effect from closest defender distance to the shot outcome. The analysis is aimed to establish whether being away from the defender, i.e. having more time to prepare and execute a shot enables the shooter to score with a higher probability. Opposite effect will be when defender is close to the shooter it doesn't leave the shooter enough time to prepare and execute the shot properly.

###Plot Result

** Table 13**: Relationship between the defender distance and accuracy

###Correlation Test
  The next analysis calculates the correlation between both variables. If the analysis included all data points for every distance, the correlation result would be as follows:
  There is a medium to strong positive correlation between the defender distance and shooting accuracy. The Pearson’s correlation coefficient for the relationship between the variables "Defender Distance" and "Accuracy" was r(297) = .59, p < .01, with a level of significance of p = .05.

###Correlation Test Result  
  When analysing the relationship between distance and the accuracy for the range of 0 to 8 feet distance the result of the correlation is different:  
  There was medium negative correlation between the defender distance and shooting accuracy in this range. The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = -0.40, p < .01, with the level of significance of p = .05.
```{r, include=FALSE}
#populate data from data frame
distanceZerotoEight <- c()
accuracyDist2 <- c()
for (i in 0:80){
  distanceZerotoEight <- c(distanceZerotoEight, distance_level[i, "Var1"])
  accuracyDist2 <- c(accuracyDist2, distance_level[i, "Accuracy"])
}

***

###Shooting Distance vs Shot Accuracy
####Introduction

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Plotting the result
library(ggplot2)
library(scales)
ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) + 
  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) + 
  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
  scale_x_discrete("Shooting Distance", limits = 0:43) + 
  geom_smooth() + 
  ylab("Shooting Accuracy (%)") + 
  ggtitle("Shooting Distance vs Shooting Accuracy")
```

###Plot Result

**Table 14**: The relationship between the shooter distance ranges and accuracy

###Correlation Test Result  
  If the correlation test taken for a certain range of distance based on the diagram, the correlation result would be different as follows:
  There was low positive correlation found between the shooting distance and shooting accuracy for the range 10 to 20 feet distance. The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(99) = .11, p =.24, failing to  the level of significance (p = .05).
```{r, include=FALSE}
#populate data for shooting distance in range 10 to 20 feet
shootingDistance10to20 <- c()
shootingAccuracyDist2 <- c()
for (i in 100:200){
  shootingDistance10to20 <- c(shootingDistance10to20, shooting_distance_level[i, "Var1"])
  shootingAccuracyDist2 <- c(shootingAccuracyDist2, shooting_distance_level[i, "Accuracy"])
}

```{r, echo=FALSE}
cor.test(dfShooting10to20$shootingDistance10to20, dfShooting10to20$shootingAccuracyDist2)
```

***

##Hot Hand Theory
###Introduction

###Barplots
```{r, include=FALSE}
#Creating a data frame for the strikes of each one of the 3 players (cannot create a data frame including all of them, because each list has different length)
player1_df <- data.frame('streaks' = streak_per_player[[1]])
player2_df <- data.frame('streaks' = streak_per_player[[2]])
player3_df <- data.frame('streaks' = streak_per_player[[3]])
```

###Boxplots

###Kyrie Irving
```{r, include=FALSE}
###PLAYER 1 VS SAMPLE PLAYER###
#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
set.seed(974)
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = sum(best_players_df[1,4], best_players_df[1,5], best_players_df[1,6]), replace = TRUE, prob = c(shooting_percentage[[1]], 1 - shooting_percentage[[1]]))
simul_shooter

```{r, echo=FALSE}
#Comparing barplots of sample shooter and real shooter1
#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Kyrie Irving')
grid.arrange(g1, g2, ncol = 2)
```

### Mo Williams
```{r, include=FALSE}
###PLAYER 2 VS SAMPLE PLAYER###
#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
set.seed(572)
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = sum(best_players_df[2,4], best_players_df[2,5], best_players_df[2,6]), replace = TRUE, 
                        prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
simul_shooter

###Klay Thompson
```{r, include=FALSE}
###PLAYER 3 VS SAMPLE PLAYER###
#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter3 
set.seed(57)
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = sum(best_players_df[3,4], best_players_df[3,5], best_players_df[3,6]), replace = TRUE, 
                        prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
simul_shooter

***
##Conclusion

@@

***
##References

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```
#####Report produced by: Justin Leiendecker, Alexander Romanenko, Artemis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
#####Date submitted: 16/10/2016

---

#Introduction

@@

**1. Team related**  
  1.1 Home Advantage analysis  
  1.2 Effect of rest on teams' performance

**2. Player related**  
  2.1 Effects on Accuracy  
    2.1.1 Fatigue Effect (Quarter Accuracy Rate) & Shot Clock Pressure     
    2.1.2 Defender Proximity Analysis
    2.1.3 Shooting Distance vs Shot Accuracy
  2.2 Hot Hand Theory  
  2.3 Game result impact by field goal attempts

##Dataset Description:

@@

##Dataset Strengths

@@

##Dataset Limitations  
  There are several limitations of the dataset:  
    1. The dataset consists only 1 year NBA season stats (2014-2015)  
    2. The dataset comprise only the data for regular season matches  
    3. The playoff and other matches during the year are not included
    4. The dataset does not include information on free throw shots
    5. The dataset does not include final scores of the games

@@

---

## 1.1 Home Advantage analysis
###Introduction

@@

```{r, echo=FALSE, width=50}
#plot showing home and away win ratio per team
ggplot(advdf, aes(x = advdf$WINNER)) + geom_bar(aes()) + ylab("game count") + xlab("winner")
```

@@

```{r, echo=FALSE, width=120}
#plot showing home and away win ratio per team
ggplot(advdf3, aes(x = TEAM,)) + geom_point(aes(y = WIN_PCT_HOME, col = "wins HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "wins  AWAY")) + ylab("Win ratio") + xlab("Team Names") 
```

@@

---

## 1.2 Effect of rest on teams' performance

@@

@@

---

## 2.1 Effects on accuracy

### 2.1.1 Fatigue Effect (Quarter Accuracy Rate) & Shot Clock Pressure

@@

graph2 <- ggplot(NULL) + geom_smooth(data = df.acc.secondsplayers3, aes(x = seconds.left, y = acc.top, color = "Top Players")) + geom_smooth(data = acc, aes(x = seconds.left, y = accuracy, color = "Average"))
graph2 + labs(title = "Accuracy per second on shot clock left for Top Players vs Overall Average") + xlab("Seconds left on shot clock") + ylab("Accuracy")

@@

---

###2.1.2 Closest Defender Distance vs Shot Accuracy
###Introduction  
This analysis focuses on whether there is a significant effect from closest defender distance to the shot outcome. The analysis is aimed to establish whether being away from the defender, i.e. having more time to prepare and execute a shot enables the shooter to score with a higher probability. Opposite effect will be when defender is close to the shooter it doesn't leave the shooter enough time to prepare and execute the shot properly.

** Table 9**: Relationship between the defender distance and accuracy

####Correlation Test
  The next analysis calculates the correlation between both variables. If the analysis included all data points for every distance, the correlation result would be as follows:
  There is a medium to strong positive correlation between the defender distance and shooting accuracy. The Pearson’s correlation coefficient for the relationship between the variables "Defender Distance" and "Accuracy" was r(297) = .59, p < .01, with a level of significance of p = .05.

####Correlation Test Result  
  When analysing the relationship between distance and the accuracy for the range of 0 to 8 feet distance the result of the correlation is different:  
  There was medium negative correlation between the defender distance and shooting accuracy in this range. The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = -0.40, p < .01, with the level of significance of p = .05.
```{r, include=FALSE}
#populate data from data frame
distanceZerotoEight <- c()
accuracyDist2 <- c()
for (i in 0:80){
  distanceZerotoEight <- c(distanceZerotoEight, distance_level[i, "Var1"])
  accuracyDist2 <- c(accuracyDist2, distance_level[i, "Accuracy"])
}

---

###2.1.3 Shooting Distance vs Shot Accuracy
###Introduction

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Plotting the result
library(ggplot2)
library(scales)
ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) + 
  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) + 
  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
  scale_x_discrete("Shooting Distance", limits = 0:43) + 
  geom_smooth() + 
  ylab("Shooting Accuracy (%)") + 
  ggtitle("Shooting Distance vs Shooting Accuracy")
```
###Results

**Table 10**: The relationship between the shooter distance ranges and accuracy

####Correlation Test Result  
  If the correlation test taken for a certain range of distance based on the diagram, the correlation result would be different as follows:
  There was low positive correlation found between the shooting distance and shooting accuracy for the range 10 to 20 feet distance. The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(99) = .11, p =.24, failing to  the level of significance (p = .05).
```{r, include=FALSE}
#populate data for shooting distance in range 10 to 20 feet
shootingDistance10to20 <- c()
shootingAccuracyDist2 <- c()
for (i in 100:200){
  shootingDistance10to20 <- c(shootingDistance10to20, shooting_distance_level[i, "Var1"])
  shootingAccuracyDist2 <- c(shootingAccuracyDist2, shooting_distance_level[i, "Accuracy"])
}

```{r, include=FALSE}
cor.test(dfShooting10to20$shootingDistance10to20, dfShooting10to20$shootingAccuracyDist2)
```

---

## 2.2 Hot Hand Theory
###Introduction

###Results
####Barplots
```{r, include=FALSE}
#Creating a data frame for the strikes of each one of the 3 players (cannot create a data frame including all of them, because each list has different length)
player1_df <- data.frame('streaks' = streak_per_player[[1]])
player2_df <- data.frame('streaks' = streak_per_player[[2]])
player3_df <- data.frame('streaks' = streak_per_player[[3]])
```

####Boxplots

####Kyrie Irving
```{r, include=FALSE}
###PLAYER 1 VS SAMPLE PLAYER###
#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
set.seed(974)
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = sum(best_players_df[1,4], best_players_df[1,5], best_players_df[1,6]), replace = TRUE, prob = c(shooting_percentage[[1]], 1 - shooting_percentage[[1]]))
simul_shooter

```{r, echo=FALSE}
#Comparing barplots of sample shooter and real shooter1
#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Kyrie Irving')
```

#### Mo Williams
```{r, include=FALSE}
###PLAYER 2 VS SAMPLE PLAYER###
#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
set.seed(572)
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = sum(best_players_df[2,4], best_players_df[2,5], best_players_df[2,6]), replace = TRUE, 
                        prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
simul_shooter

####Klay Thompson
```{r, include=FALSE}
###PLAYER 3 VS SAMPLE PLAYER###
#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter3 
set.seed(57)
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = sum(best_players_df[3,4], best_players_df[3,5], best_players_df[3,6]), replace = TRUE, 
                        prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
simul_shooter

---

##2.3 Game result impact by field goal attempt

@@

#### Impact of Top Scorers

@@

```
**Table 11**: Result for the logistic regression analysis for FGA and FG% of 15 players

@@

**Table 12**: Result for the anova for FGA and FG% for 15 players

@@

**Table 13**: Result for the logistic regression analysis for FGA and FG% of the 3 special players

@@

**Table 14**: Result for the anova for FGA and FG% for the 3 special players

@@

---

##Conclusion

@@

---

##References
